---
title: "Bioinformatics_Exam_Q15"
author: "Galilea Guerrrero PID: A59010531"
date: '2022-07-13'
output: html_document
---

Import data from the Human Services (CHHS) open data site:
https://data.chhs.ca.gov/dataset/covid-19-variant-data and loaded required packages.
```{r}
data <- read.csv("covid19_variants.csv")

library(ggplot2)

library(lubridate)

library(tidyverse)
```
Filtered data set to remove "Total" and "Other" data points from variants.
```{r}
data2 <- filter(data, variant_name != "Total")
head(data2)

data3 <- filter(data2, variant_name != "Other")
head(data3) 
```

Specify dates as a date variable.
```{r}
data3$date <- ymd(data3$date)
```


Make a line plot with variants over time. 
```{r}
plot <- ggplot(data3) + aes(date, percentage, group=variant_name) + 
  geom_line(aes(color=variant_name)) + 
  ggtitle("COVID-19 Variants in California")+
  labs(x="", y="Percentage of Sequenced Specimens") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y")+
  scale_color_discrete(name="")

plot + theme_light() + theme(axis.text.x = element_text(size=6, vjust=0.5, angle=45)) 
```

Final plot represents the prominence of COVID variants in California as a percentage of sequenced specimens over time (January 2021-July 2022). 

Alpha and Delta variants appear to become prominent in the population rapidly and for a prolonged period of time. 

